Discussion Draft

### 2020-04-04 doc/Discussion.rmd
spell check

Under **spindle based model** we hypothesize that the sexual dimorphism in the recombination landsacpes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I. <1> the irreversable process of the metaphase to anaphase transistion is initiated by the protyltic decay of the sister cohesion connecting homologs --along the tetrad arms (cite).

<2> The number and placement of crossovers alter the distribution of sister cohesion and the 'chromosome stucture' once kinetochores attached to MT-spindle and tension applied.  For single acrocentric chromosomes the sister cohesion distal to the crossover connects the homologs (when aligned on metaphase plate). This is known as cruciform tetrad is some systems.

This area is reduced when crossover position is shifted towards the telomere (ends of the chromsome).

The male crossover telomere bias could be adaptive in order to synchornize (or minimize) the timing of homologs sepearting at metaphase I.

<3> Conserved feature of gametogenesis -- is the length of timing.
faster spermatogenesis -- may select for synchronize the separation homologs or faster seperate after alignment at metaphase. While in oogenesis, the longer timing and arrests may select for overall more sister cohesion (a different homolog strucutres)

@@

Our results from MSM and PWD show that the sc length compaction is independant (of the direction of heterochiasmy. Two strains have have more crossovers in males, but females still have longer sc axis. **This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape / heterochiasmy**

@@

Interference is a fundamental aspect of the recombination landscape. -- genome-wide recombination rate and chromaint compaction

@@

Our results support an opposite pattern predicted from the 'logical model', a positive correlation between interference strength and genome wide recombination rates; The within sex comparison of two breeds of cattle with different genome wide recombintion rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].

>>>>

Under **spindle based model** we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I. <1> the irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (referred to as along the tetrad arms (cite)).

<2> The number and placement of crossovers alter the distribution of sister cohesion and the 'chromosome structure' which is a consequence of kinetochores attaching to the MT-spindle and tension applied. Single crossover bivalents form what is known as cruciform tetrad in some systems (figure x). This area is reduced when crossover positions are shifted towards the telomere ends of the chromosome. The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.

<3> Conserved feature of gametogenesis -- is the length of timing of the cell cycle (difference between spermat.
faster spermatogenesis -- may select for synchronize the separation homologs or faster separate after alignment at metaphase. While in oogenesis, the longer timing and arrests may select for overall more sister cohesion (a different homolog structures)

@@

Our results from MSM and PWD show that the SC length compaction is independent (of the direction of heterochiasmy. Two strains have have more crossovers in males, but females still have longer SC axis. **This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape / heterochiasmy**

@@

Interference is a fundamental aspect of the recombination landscape. -- genome-wide recombination rate and chromatin compaction

@@

Our results support an opposite pattern predicted from the 'logical model', a positive correlation between interference strength and genome wide recombination rates; The within sex comparison of two breeds of cattle with different genome wide recombination rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].
